\dppTitle

### 2014-06-16 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Making some edits to dpp ms in proofing stage.
Adding Table S1 to supporting information that give an example of the
divergence model sample space when Y=4 taxa.

Also, making some edits to the method section where the sample space of
ordered/unordered divergence models is introduced. Making the language
in this section more exact and introducing my loose use of "ordered" and
"unordered" subsequent to this section.

\begin{linenomath}
As for the prior term $p(\divTimeIndexVector)$, the total sample space of
ordered realizations of \divTimeIndexVector is all the possible partitions of
\npairs{} elements into 1 to \npairs{} categories.
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible numbers of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell Number \cite{Bell1934}.
The original \msb model only samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
    OeisPartitionTriangle,Malenfant2011}.
I denote the set of all possible integer partitions of \npairs{} as
\integerPartitionSet{\npairs{}} and the length of that set as
\integerPartitionNum{\npairs{}}.
The advantages, disadvantages, and justification of ignoring the order
of \divTimeIndexVector is discussed in detail below.
\end{linenomath}

@@

\subsection{Assessing model-choice behavior and robustness}
Following the simulation-based approach of Oaks et al.\ \cite{Oaks2012}, I
characterize the behavior of several models under the ideal conditions where
the data are generated from parameters drawn from the same prior distributions
used for analysis (i.e., the prior is correct).
I selected the following four model priors for these analyses
(Table~\ref{tabPriors}).
\begin{enumerate}
    \item The \modelOld model represents the original \msb implementation with
        the U-shaped prior on unordered divergence models and uniform priors on
        divergence-time and demographic parameters; $\divTimeIndexVector \sim
        \priorOld$, $\divTime{} \sim U(0,10)$, $\ancestralTheta{} \sim U(0,
        0.05)$, and $\descendantThetaMean{} \sim U(0, 0.05)$.
    \item The \modelUshaped model with the U-shaped prior of \msb on unordered
        divergence models, but with exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorOld$,
        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
    \item The \modelUniform model with a uniform prior over unordered
        divergence models and exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorUniform$,
        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
    \item The \modelDPP model with a Dirichlet-process prior on ordered
        divergence models and exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorDPP{\sim
        Gamma(2,2)}$, $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{}
        \sim Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$,
        and $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
\end{enumerate}
I selected the exponential prior on divergence time used in models \modelDPP,
\modelUniform, and \modelUshaped to have the same variance as the uniform prior
in model \modelOld.
I selected the exponential prior on population size used in models \modelDPP,
\modelUniform, and \modelUniform to have the same mean as the uniform prior in
model \modelOld, so that all four models have the same \myTheta{C} and thus the
same units of time.
All of the models were the same in other respects, with three free \myTheta{}
parameters for each population pair, two uniformly distributed ($beta(1,1)$)
\bottleScalar{}{} parameters per pair, no migration, no recombination, and
resorting of taxon-specific summary statistics by $\pi_b$ (i.e., sampling
unordered divergence models).
For all simulations, I used a data structure of eight population pairs, with a
single 1000 base-pair locus sampled from 10 individuals from each population.

>>>>

\begin{linenomath}
\change{
As for the prior term $p(\divTimeIndexVector)$, the total sample space of
\divTimeIndexVector is all the possible partitionings of the \npairs{} pairs of
populations into 1 to \npairs{} divergence-time classes, where each
partitioning consists of non-overlapping and non-empty subsets whose union is
the \npairs{} pairs;
Hereinafter, I refer to these partitionings as ``ordered'' divergence models or
partitions.
}
The total number of possible partitions is a sum of the Stirling numbers of
the second kind over all possible numbers of categories \divTimeNum
\begin{equation}
    B_{\npairs{}}=\sum_{\divTimeNum=1}^{\npairs{}} \left[
    \frac{1}{\divTimeNum!} \sum_{j=0}^{\divTimeNum-1} (-1)^{j}
    \binom{\divTimeNum}{j} (\divTimeNum-j)^{\npairs{}} \right],
    \label{eq:bell}
\end{equation}
which is the Bell number \cite{Bell1934}.
The original \msb model samples over the unordered realizations of
\divTimeIndexVector, such that the sample space is reduced to all the possible
integer partitions of \npairs{} \cite{Oaks2012,Huang2011,OeisPartitionNumber,
OeisPartitionTriangle,Malenfant2011} (Table~S\ref{tabSampleSpace}).
\change{
I denote the set of all possible integer partitions of the \npairs{} pairs of
populations as \integerPartitionSet{\npairs{}} and the length of that set as
\integerPartitionNum{\npairs{}}, and I hereinafter refer to these integer
partitions as ``unordered'' divergence models or partitions.}
The advantages, disadvantages, and justification of ignoring the order
of \divTimeIndexVector is discussed in detail below.
\end{linenomath}

@@

\subsection{Assessing model-choice behavior and robustness}
Following the simulation-based approach of Oaks et al.\ \cite{Oaks2012}, I
characterize the behavior of several models under the ideal conditions where
the data are generated from parameters drawn from the same prior distributions
used for analysis (i.e., the prior is correct).
I selected the following four model priors for these analyses
(Table~\ref{tabPriors}).
\begin{enumerate}
    \item The \modelOld model represents the original \msb implementation with
        the U-shaped prior on unordered divergence models and uniform priors on
        divergence-time and demographic parameters; $\divTimeIndexVector \sim
        \priorOld$, $\divTime{} \sim U(0,10)$, $\ancestralTheta{} \sim U(0,
        0.05)$, and $\descendantThetaMean{} \sim U(0, 0.05)$.
    \item The \modelUshaped model with the U-shaped prior of \msb on unordered
        divergence models, but with exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorOld$,
        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
    \item The \modelUniform model with a uniform prior over unordered
        divergence models and exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorUniform$,
        $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{} \sim
        Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$, and
        $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
    \item The \modelDPP model with a Dirichlet-process prior on ordered
        divergence models and exponential priors on divergence-time and
        demographic parameters; $\divTimeIndexVector \sim \priorDPP{\sim
        Gamma(2,2)}$, $\divTime{} \sim Exp(mean=2.887)$, $\ancestralTheta{}
        \sim Exp(mean=0.025)$, $\descendantTheta{1}{} \sim Exp(mean=0.025)$,
        and $\descendantTheta{2}{} \sim Exp(mean=0.025)$.
\end{enumerate}
I selected the exponential prior on divergence time used in models \modelDPP,
\modelUniform, and \modelUshaped to have the same variance as the uniform prior
in model \modelOld.
I selected the exponential prior on population size used in models \modelDPP,
\modelUniform, and \modelUniform to have the same mean as the uniform prior in
model \modelOld, so that all four models have the same \myTheta{C} and thus the
same units of time.
All of the models were the same in other respects, with three free \myTheta{}
parameters for each population pair, two uniformly distributed ($beta(1,1)$)
\bottleScalar{}{} parameters per pair, no migration, no recombination, and
re-sorting of taxon-specific summary statistics by $\pi_b$ (i.e., sampling
unordered divergence models).
For all simulations, I used a data structure of eight population pairs, with a
single 1000 base-pair locus sampled from 10 individuals from each population.
